---
title: 02 Lines
execute: 
  cache: true
  freeze: auto
knitr:
  opts_chunk: 
    warning: false
    message: false
filters:
  - lightbox
lightbox: auto
---

A map with lines.

## Libraries
```{r}
#| results: false
#| code-fold: true
library(sf)
library(tmap)
library(dplyr)
library(ggplot2)
```

## Data wrangling
```{r}
#| results: false
#| code-fold: true
nz <- st_read("data/nz.gpkg") %>%
  st_cast("MULTILINESTRING") %>%
  mutate(What = "Shore",
         What = as.factor(What))
```

## The maps
But first... a rabbithole.

Something this map got me looking at a bit more closely than ever before was exactly what is going on with the `linewidth` setting in `ggplot2`, which, like a lot of things you've learned by osmosis, rather than systematically, doesn't seem that important until it does. 

The obvious fact is that `lwd = 0.5` in `tmap` and `linewidth = 0.5` in `ggplot2` do not yield the same result. Turns out... it's complicated.  `tmap` uses 0.75 [points](https://en.wikipedia.org/wiki/Point_(typography)), and `ggplot2` uses ~0.75 [millimetres](https://en.wikipedia.org/wiki/Millimetre).

Wait, wut? 0.75 pts and _approximately_ 0.75 mm?! Yes, you read that correctly. 

To be fair to `tmap`, it's using the _R_ graphics standard, which is that `lwd` is specified in pixels, which is notionally 1/96 inch, as [laid down by CSS](https://www.w3schools.com/cssref/css_units.php). This is 'device dependent' and when it comes to getting plots out of _R_ and into image file formats can lead to all kinds of headaches, but that's another story. A point is 1/72 inch (ever since [Adobe made it so](https://en.wikipedia.org/wiki/Point_(typography)#Desktop_publishing_point)), so `lwd = 1` gives you nominal line widths of 0.75 points in `tmap`.

Meanwhile, according to [the `ggplot2` documentation](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html?q=linewidth#linewidth) (RTFM, eh?)

> Due to a historical error, the unit of linewidth is roughly 0.75 mm. Making it exactly 1 mm would change a very large number of existing plots, so we’re stuck with this mistake.

They say nothing more on the matter. Presumably, it's something to do with that 1/72 _vs_ 1/96 thing just noted, and perhaps the `approximately` is due to TeX's use of a point being 1/72.27 inches, [which is apparently more historically accurate](https://tex.stackexchange.com/questions/200934/why-does-a-tex-point-differ-from-a-desktop-publishing-point), and was the approximation settled upon, _because it is palindromic_! OK... welll maybe not really that last bit. 

Anyway, the magic number 72.27 is buried in the `ggplot2` code, where the [conversion factor 72.27 / 25.4 is baked in](https://github.com/tidyverse/ggplot2/blob/b05c6c1f61a52311809ffda775eb258b07518dab/R/geom-.r#L163). I was led to this information by [this stackoverflow post](https://stackoverflow.com/questions/47519624/how-is-the-line-width-size-defined-in-ggplot2).

The upshot of all this? To get the same line widths in `ggplot2` as in `tmap` multiply by 25.4 / 72.27. Simple really!

### `tmap`
```{r}
tm_shape(nz) + 
  tm_lines(
    col = "What",
    col.scale = tm_scale_categorical(
      values = c("blue")),
    lwd = 0.5) + 
  tm_layout(
    frame = FALSE,
    legend.frame = FALSE,
    legend.outside = TRUE)
```

### `ggplot2`
```{r}
ggplot(nz) +
  geom_sf(aes(colour = What), linewidth = 0.5 * 25.4 / 72.27) +
  scale_colour_manual(values = c("blue")) +
  theme_void()
```

