---
title: 07 Navigation
execute: 
  cache: true
  freeze: auto
knitr:
  opts_chunk: 
    warning: false
    message: false
---

## Libraries
```{r}
#| results: false
#| code-fold: true
library(sf)
library(tmap)
library(dplyr)
library(geosphere)
library(ggplot2)
library(stringr)

source("~/Documents/code/30-day-maps-2023/maps/utils.R")
```

## Data wrangling
```{r}
#| results: false
#| code-fold: true
start_lat <- -41.3
start_lon <- 174.75
end_lat <- 45
ortho_proj <- str_glue("+proj=ortho +lat_0={start_lat} +lon_0={174.75}")
bearing <- -88.5
step_length <- 5e4
last_pt <- c(start_lon, start_lat)
lox <- list()
transect <- last_pt
i <- 1
while (last_pt[2] < end_lat) {
  next_pt <- destPointRhumb(last_pt, bearing, step_length)
  if (next_pt[1] > last_pt[1]) {
    lox[[i]] <- transect
    i <- i + 1
    transect <- c(next_pt)
  } else {
    transect <- c(transect, next_pt)
  }
  last_pt <- next_pt
}
lox[[i]] <- transect

lox_sf <- lox %>%
  lapply(unlist) %>%
  lapply(matrix, ncol = 2, byrow = TRUE) %>%
  st_multilinestring() %>%
  st_sfc() %>%
  st_sf(crs = 4326) %>%
  st_cast("LINESTRING") %>%
  st_cast("POINT")

data("World")
world <- World %>%
  st_cast("POLYGON") %>%
  st_transform(ortho_proj) %>%
  dplyr::filter(!(st_is_empty(geometry)))

graticule <- get_graticule(centre = c(start_lon, start_lat)) %>%
  st_transform(ortho_proj)
```

## `tmap`
At time of writing there is a bug in `tm_title` so no title on this map. See [this issue](https://github.com/r-tmap/tmap/issues/796).

Note that I hand made the graticule using code from a sourced *R* file which trims the graticule to only extend across the mapped area. Neither `tmap` nor `ggplot2` successfully applied the graticule in this projection without errors using their native functionality.

```{r}
tm_shape(world) +
  tm_fill(fill = "#cccccc", lwd = 0) +
  tm_shape(graticule, is.main = TRUE) +
  tm_lines(col = "grey", lwd = 0.5) +
  tm_shape(lox_sf) +
  tm_dots(fill = "firebrick", size = 0.001) +
  # tm_title("    Go (just north of) west! See also Day 22") +
  tm_layout(frame = FALSE)
```

## `ggplot2`
```{r}
ggplot(world) +
  geom_sf(fill = "#cccccc", linewidth = 0) +
  geom_sf(data = graticule, colour = "grey", linewidth = 0.2) +
  geom_sf(data = lox_sf, colour = "firebrick", size = 0.01) +
  ggtitle("Go (just north of) west! See also Day 22") +
  theme_void()
```
