---
title: 13 Choropleth
# execute: 
#   cache: true
#   freeze: auto
knitr:
  opts_chunk: 
    warning: false
    message: false
---

Classic thematic map: a choropleth.

This one is where `tmap` comes into its own, although v4 in increasing the available options requires more code to make the same map compared to v3. `ggplot2` by contrast is allergic to making classic classified choropleth maps, so much so that it's hard to imagine using it in any kind of introductory teaching setting.

## Libraries
```{r}
#| results: false
#| code-fold: true
library(sf)
library(tmap)
library(dplyr)
library(ggplot2)
```

## Data wrangling
```{r}
#| results: false
#| code-fold: true
ak <- st_read("data/ak-demographics-2018.gpkg") %>% 
  slice(-1) # remove Gulf Islands
```

## `tmap`
```{r}
tm_shape(ak) + 
  tm_polygons(
    fill = "maori",
    fill.scale = tm_scale_intervals(
      n = 9, values = "YlGnBu", 
      style = "quantile"),
    fill.legend = tm_legend(
      orientation = "landscape",
      title = "% Māori"),
    col = "grey",
    lwd = 0.35
  ) +
  tm_layout(
    frame = FALSE, legend.outside = TRUE,
    legend.outside.position = "bottom",
    legend.frame = FALSE)
```

## `ggplot2`
```{r}
library(classInt)

intervals <- classIntervals(ak$maori, 9, "quantile", digits = 1)
brks <- round(intervals$brks, 1)
lbls = paste0(brks[-10], rep(" to ", 9), brks[-1])
bb <- st_bbox(ak)

ggplot(ak) + 
  geom_sf(aes(fill = maori), colour = "grey", linewidth = 0.2) +
  scale_fill_fermenter(
    palette = "YlGnBu",
    direction = 1,
    breaks = brks[-10],
    labels = lbls,
    name = "% Māori",
    guide = guide_legend(
      keywidth = unit(15, units = "mm"),
      keyheight = unit(3, units = "mm"),
      label.position = "bottom",
      title.position = "top",
      nrow = 1)) +
  coord_sf(ylim = c(bb[2] - 10000, bb[4]), clip = "off") +
  theme_void() +
  theme(legend.position = c(0.5, 0.1))

```
