---
title: 20 Outdoors
execute: 
  cache: true
  freeze: auto
knitr:
  opts_chunk: 
    warning: false
    message: false
filters:
  - lightbox
lightbox: auto
---

Map of mountains, trails, or something completely different.

As [previously noted](18-atmosphere.qmd) much of the outdoors in Wellington is dominated by wind. 

That's why I've used  a pale blue for all of the space outside the buildings in these maps. If it was a warmer time of year, that blue might be a bit less watery and cold.

The visual inspiration is _machine space_

> Horvath RJ. [Machine space](https://www.csun.edu/~rdavids/350fall08/350readings/Horvath%20Machine%20space.pdf). Geographical Review **64**(2) 167-188.

but inverted. Of course, a lot of the 'open' space here is also machine space (given over to cars), but still, it's fun to look a the built environment in this way.

The buildings are from [Te Toitū Whenua - Land Information New Zealand](https://data.linz.govt.nz/layer/101290-nz-building-outlines/). Tracks are [from Wellington City Council](https://catalogue.data.govt.nz/dataset/wcc-tracks4).

But the point is, that the buildings aren't the spatial dataset here: `outdoors.gpkg` is a rectangle of the extent of the map with holes where the buildings (the 'indoors') are. So it's a 'outdoors' polygon. You can't really tell that from the map, but the white areas are the background colour showing through the holes.

## Libraries
```{r}
#| results: false
#| code-fold: true
library(sf)
library(tmap)
library(ggplot2)
```

## Data wrangling
```{r}
#| results: false
#| code-fold: true

outdoors <- st_read("data/outdoors.gpkg")
tracks <- st_read("data/WCC_Tracks.geojson")
coast <- st_read("data/welly-outdoors-coast.gpkg")
```

## The maps
### `tmap`
```{r fig.width=10, fig.height=10}
tm_shape(outdoors) +
  tm_fill(fill = "#aaddff") +
  tm_shape(tracks) +
  tm_lines(col = "#66aacc", lwd = 0.2) +
  tm_shape(coast) +
  tm_lines(col = "#aaddff", lwd = 0.5) +
  tm_scalebar() +
  tm_title("Te Whanganui-a-Tara outdoors",
    position = tm_pos_out(
      cell.h = "center", cell.v = "top",
      pos.h = "left", pos.v = "top")) +
  tm_layout(frame = FALSE, inner.margins = rep(0, 4))
```

## `ggplot2`
```{r fig.width=10, fig.height=10}
library(ggspatial)

ggplot() +
  geom_sf(data = outdoors, fill = "#aaddff", linewidth = 0) +
  geom_sf(data = tracks, colour = "#66aacc", 
          linewidth = 0.2 * 25.4 / 72.27) +
  geom_sf(data = coast, colour = "#aaddff",
          linewidth = 0.5 * 25.4 / 72.27) +
  coord_sf(xlim = st_bbox(outdoors)[c(1, 3)],
           ylim = st_bbox(outdoors)[c(2, 4)],
           expand = FALSE) +
  annotation_scale() +
  ggtitle("Te Whanganui-a-Tara outdoors") +
  theme_void()
```