---
title: "26 Minimal"
execute: 
  cache: true
  freeze: auto
knitr:
  opts_chunk: 
    warning: false
    message: false
lightbox: auto
---

Less is more.

Probably not really in keeping with the theme for this one, but I once claimed (in [_Geographical Information Analysis_](https://www.wiley.com/en-us/Geographic+Information+Analysis%2C+2nd+Edition-p-9780470288573)) that a recognisable map of Australia could be formed from only 9 points (sorry Tasmania...). For this theme, I thought I'd take a look at what a minimal New Zealand might look like. Funny thing is, I think you need more points for a recognisable New Zealand.

The source is this [Statistics NZ SA2 geographic boundaries dataset](https://datafinder.stats.govt.nz/layer/92210-statistical-area-1-2018-generalised/) selecting only the mainland and island areas, and dissolving them into a single multipolygon. This has around 880,000 vertices. There are a lot of small islands in this dataset (over 1700) and these are rapidly sacrificed by the generalisation process. It is worth noting that _R_ hangs when trying to plot this layer ungeneralised, at least on my laptop (even with 64GB of RAM).

Anyway, to tackle this, I used the [`rmapshaper` package](http://www.mapshaper.org/) which can do topological simplification of polygons. You specify a `keep` parameter for the fraction of vertices to retain in the simplified geometries. For reasons of binary division, I started from 5.12% (the `ms_simplify` function defaults to 5%) and progressed in steps of 50%, all the way down to retaining 0.0025%. At 1 vertex in 40,000 or so of the original geometry we get quite a nice 25 vertex 'minimal' New Zealand. After that things start to go a bit pear-shaped (almost literally) as I show at the end of this page.

Stewart Island survives until 0.01% a step more than Lake Taupō, while the Chathams drop out a step before that.

## Libraries
```{r}
#| results: false
#| code-fold: true
library(sf)
library(rmapshaper)
library(tmap)
library(ggplot2)
library(dplyr)
```

## Data wrangling
The two main pain points here are, first, that it is much faster to iteratively apply `keep = 0.5` to each step, than to start with the full map and keep 5.12%, then 2.56%, then 1.28% and so on; and second, that getting the factor order right is fiddly. I wanted the facet plots to proceed from less generalised (higher % values) to more, and that requires care in setting up the factor.

```{r}
#| results: false
#| code-fold: true
nz <- st_read("data/nz-large.gpkg")

start_fraction <- 0.0512
num_steps <- 12

geoms <- c(nz$geom %>% ms_simplify(start_fraction))
for (i in 2:num_steps) {
  geoms <- c(geoms,
             geoms[i-1] %>% ms_simplify(keep = 0.5))
}
percents <- paste(
  start_fraction * 100 * (0.5 ^ (1:num_steps - 1)), "%", sep = "")

nzs <- 
  data.frame(
    Percent = ordered(percents, levels = percents), 
    geom = geoms %>% st_sfc()) %>%
  st_as_sf()
```

## The maps
### `tmap`
This case lets us see how `tm_facets` works. Pretty simple really. `free.coords = FALSE` is necessary to make sure the panels are all scaled the same. 

```{r fig.width=10, fig.height=8.5}
tm_shape(nzs) + 
  tm_polygons(col = "black") +
  tm_layout(main.title = "Generalisation % of vertices retained") +
  tm_facets(by = "Percent", ncol = 4, free.coords = FALSE)
```

### `ggplot2`
`facet_wrap` handles this very nicely.

```{r fig.width=10, fig.height=8.5}
ggplot(nzs) +
  geom_sf(linewidth = 25.4 / 72.27) +
  facet_wrap( ~ Percent, ncol = 4) + 
  labs(title = "Generalisation % of vertices retained") +
  theme_void() +
  theme(panel.border = element_rect(fill = NA))
```

So... here's a pretty recognisable minimal map of New Zealand, with 25 points:

```{r}
#| classes: preview-image
plot(geoms[12])
```

One more step and it gets pretty bad:

```{r}
plot(geoms[12] %>% 
       ms_simplify(keep = 0.5))
```

One more gets us an unexpectedly recognisable 10 vertices:

```{r}
plot(geoms[12] %>% 
       ms_simplify(keep = 0.5) %>% 
       ms_simplify(keep = 0.5))
```

After that it really doesn't work:

```{r}
plot(geoms[12] %>% 
       ms_simplify(keep = 0.5) %>% 
       ms_simplify(keep = 0.5) %>% 
       ms_simplify(keep = 0.5))
```
